### 2023-03-16 paper.tex
write methods

\subsection{Re-analysis of published scBS datasets}
To re-analyze scBS data sets by \citet{kremer_scnmt} and \citet{luo2017single}, the genome was first annotated with genomic intervals that were either obtained by tiling the genome into non-overlapping 100~kb tiles, or by acquiring VMRs with \texttt{scbs scan} using the current default options (bandwidth $=2000$, stepsize $=10$, variance threshold $= 0.02$).
VMRs with sequencing coverage in less than 
% WIP
For all intervals (VMRs and 100~kb tiles), the mean methylation fractions and the shrunken means of the residuals were obtained with \texttt{scbs matrix}.

@@

\subsection{Nearest neighbor count score} \label{methods:score}
To assess performance of our methods, we processed mouse CpG methylation data from \citet{luo2017single} with several method combinations and scored each result.
The score evaluates the ability of each method combination to separate cell types in 15-dimensional PCA space, based on their CpG methylation.
Each score varies between 0 and 1, where higher scores reflect better separation.

@@

\section{Supplementary Figures}

@@

\begin{figure*}[h]
    \begin{center}
        \includegraphics[width=0.8\columnwidth]{figures/SFig_benchmark.png}
    \end{center}
    \caption{\small \textbf{Benchmark of our methods on single-cell methylomes of neural subtypes of the murine frontal cortex}.\\
    Our methods were benchmarked on single-cell methylomes from \citet{luo2017single} as described in Fig.~\ref{fig:score}.
    }
    \label{fig:score_luo}
\end{figure*}

\begin{figure*}[h]
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/SFig_parameter_sweep.png}
    \end{center}
    \caption{\small \textbf{Effect of VMR detection parameters on separation of cell types}.\\
    \textbf{(A-B)} Mean neighbor scores obtained after analyzing single-cell methylomes from \citet{kremer_scnmt} (\textbf{A}) or \citet{luo2017single} (\textbf{B}) with our proposed methods.
    VMRs were detected with \texttt{scbs scan} using various sliding window bandwidths or variance thresholds, and on sub-samples of the full data sets.
    }
    \label{fig:sweep}
\end{figure*}

>>>>

\subsection{Re-analysis of published scBS datasets}
To re-analyze scBS data from \citet{kremer_scnmt}, single-cell CpG methylation reports were first stored with \texttt{scbs prepare} and then smoothed with \texttt{scbs smooth} using the default bandwidth of 1000~bp.
The mm10 mouse genome was then annotated with genomic intervals that were either obtained by tiling the genome into non-overlapping 100~kb tiles, or by acquiring VMRs with \texttt{scbs scan} using the current default options (bandwidth $=2000$, stepsize $=10$, variance threshold $= 0.02$, minimum cell requirement $=6$).
Methylation fractions, commonly represented as percentages, and the shrunken means of residuals were quantified with \texttt{scbs matrix} for both VMRs and tiles.
The same analysis was repeated for scBS data from \citet{luo2017single}, and for smaller data sets generated by randomly sub-sampling cells separately from both data sets.
For each data set, we thus obtained four methylation matrices:
either 100~kb tiles or VMRs, quantified either by simple averaging or with the shrunken mean of the residuals.
All four matrices were then subjected to PCA, followed by UMAP with parameters min\_dist $=0.2$, init $=$ spca, and n\_neighbors $=\frac{\sqrt{n}}{1.5}$ where $n$ is the total cell number, to adapt to data sets of different cell numbers.

\subsection{Mean neighbor score} \label{methods:score}
To assess performance of our methods, we employed a score

we processed mouse CpG methylation data from \citet{luo2017single} with several method combinations and scored each result.
The score evaluates the ability of each method combination to separate cell types in 15-dimensional PCA space, based on their CpG methylation.
Each score varies between 0 and 1, where higher scores reflect better separation.

% WIP
For all intervals (VMRs and 100~kb tiles), the mean methylation fractions and the shrunken means of the residuals were obtained with \texttt{scbs matrix}.

@@

@@

\newpage
\section{Supplementary Figures}

@@

\begin{figure*}
    \begin{center}
        \includegraphics[width=0.8\columnwidth]{figures/SFig_benchmark.png}
    \end{center}
    \caption{\small \textbf{Benchmark of our methods on single-cell methylomes of neural subtypes of the murine frontal cortex}.\\
    Our methods were benchmarked on single-cell methylomes from \citet{luo2017single} as described in Fig.~\ref{fig:score}.
    }
    \label{fig:score_luo}
\end{figure*}

\begin{figure*}
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/SFig_parameter_sweep.png}
    \end{center}
    \caption{\small \textbf{Effect of VMR detection parameters on separation of cell types}.\\
    \textbf{(A-B)} Mean neighbor scores obtained after analyzing single-cell methylomes from \citet{kremer_scnmt} (\textbf{A}) or \citet{luo2017single} (\textbf{B}) with our proposed methods.
    VMRs were detected with \texttt{scbs scan} using various sliding window bandwidths or variance thresholds, and on sub-samples of the full data sets.
    }
    \label{fig:sweep}
\end{figure*}
